---
title: "CrossSpecies Plot"
output: html_document
---
#Load Necessary Packages
```{r, warning=F, message=F}
#If ggtree has not previously been installed, remove the "#" on the line below and use BiocManager to download the package. Note if you do not "biocmanaged" package installed, you will need that as well.
#BiocManager::install("ggtree")

#Load the following packages to run the res of the script.
library(tidyverse)
library(viridis)
library(Rmisc)
library(phylogram)
library(ggtree)
library(ggplot2)
library(nlme)
library(arsenal)
```

## Introduction Plot: Bias in IGF Publication 

```{r}
pub=read.csv("Timeline_IGFs_pub.csv")
pub=subset(pub, Year != "2021")

ggplot() + 
 geom_line(data=pub, aes(x = Year, y = IGF1), color= "orange2", fill = "orange2", size = 1,  alpha=1) +
  geom_line(data=pub, aes(x = Year, y = IGF2), color= "skyblue4", fill="skyblue4",  size = 1,  alpha=1) +
  geom_line(data=pub, aes(x = Year, y = Post.natal_no.cancer_IGF1), color= "orange2", size = 1, linetype= "twodash") +
  geom_line(data=pub, aes(x = Year, y = Post.natal_no.cancer_IGF2), color= "skyblue4", size = 1, linetype= "twodash") +
 geom_area(data=pub, aes(x = Year, y = Non.model_IGF1), color= "orange2", fill="orange2", size = 1, alpha=0.4, linetype= "solid") +
  geom_area (data=pub, aes(x = Year, y = Non.model_IGF2), color= "skyblue4", fill="skyblue4", size = 1, alpha=0.4, linetype= "solid")  +
  ylab("Number of Publications") +
  xlab ("Year") +
theme(axis.title = element_text(face= "bold", size= 12))


```


##Data Curation of RNAseq Data
```{R}

# Upload dataset in CSV format
data2 <- read.csv("MetaData_Counts_Cleaned.csv")
#Get a count of the total number of unique species that will be included in the data set.
count=count(unique(data2$Organism))

#aggregate data to get average IGF1 for each species (combining all individuals of the same species)
IGF1=summarySE(data = data2, measurevar="IGF1_Proportion", groupvars = c("Organism", "Common.Name", "Category", "Subcategory", "Phylogenetic_Cat"), na.rm = FALSE, conf.interval = 0.95, .drop = TRUE)

#aggregate data to get average IGF2 for each species (combining all individuals of the same species)
IGF2=summarySE(data = data2, measurevar="IGF2_Proportion", groupvars = c("Organism", "Common.Name", "Category", "Subcategory", "Phylogenetic_Cat"), na.rm = FALSE, conf.interval = 0.95, .drop = TRUE)

#Merge IGF1 and IGF2 files together by the organisms common name.
data=merge(x = IGF1, y = IGF2, by = "Organism")

#Make new data frame including only the common name, organism, subcategory for organism organization, IGF1 value, and IGF2 value
data=data.frame(data$Common.Name.x, data$Organism, data$Category.x, data$Subcategory.x, data$Phylogenetic_Cat.x, data$IGF1_Proportion, data$IGF2_Proportion)

#rename columns for clarity following merge
names(data)[names(data) == "data.Category.x"] <- "Category"
names(data)[names(data) == "data.Common.Name.x"] <- "individual"
names(data)[names(data) == "data.Subcategory.x"] <- "group"
names(data)[names(data) == "data.IGF1_Proportion"] <- "IGF1.plot"
names(data)[names(data) == "data.IGF2_Proportion"] <- "IGF2.plot"
names(data)[names(data) == "data.Organism"] <- "Organism"
names(data)[names(data) == "data.Phylogenetic_Cat.x"] <- "Phylogenetic_Cat"

#take another count to make sure you did not lose data in the merging process
count2=count(unique(data$Organism))
#compare data frames to see what has changed
summary(comparedf(count, count2, by = "x"))

#Write file to CSV to preserve data set used in all following analyses
write.csv(data,"Concat_Final_Dat2.csv", row.names = FALSE)

#Format data frame for use in Proportions Plot. Rename columns, and then reformat to have IGF1.plot and IGF2.plot combine into a single column by species. The new column will be called "Value". This is used for proportional barplots
data3=data.frame(data$individual,data$group,data$IGF1.plot, data$IGF2.plot)
names(data3)[names(data3) == "data.individual"] <- "individual"
names(data3)[names(data3) == "data.group"] <- "group"
names(data3)[names(data3) == "data.IGF1.plot"] <- "IGF1.plot"
names(data3)[names(data3) == "data.IGF2.plot"] <- "IGF2.plot"
names(data)[names(data) == "data.Organism"] <- "Organism"

data3=data3[complete.cases(data3), ]
data3 <- data3 %>% gather(key = "observation", value="value", -c(1,2)) 
```


```{R}
library(janitor)

#stacked barplot

write.csv(data3, "datafile3.csv",row.names = FALSE)

#This file needs to be made with the adjusted values based on Randy's new Tree.
p3ID=read.csv("datafile3_plotID.csv")


  p4=ggplot(p3ID, aes(fill=observation, y=Percentage, x=plotID)) +
  geom_bar(position="stack", stat="identity") +  scale_fill_manual(values=c("orange2", "skyblue4")) +
coord_flip() +
  scale_x_discrete(lim=rev) +
 theme(axis.text.y = element_blank(), axis.title.y = element_blank())

ggsave(p4, file="phylo_barplot.nc.png", width=3, height=9, dpi=600)


```

##Violin plots by Species on qPCR data
```{r}
#Read in data and format data/subset data
sp=read.csv("Species_Combined_edited.csv")
#Set ages to order chronologically
sp$Age = factor(sp$Age, levels=c('Embryonic','Juvenile','Adult'))
#Set custom dodge location for lining us data with grouped barplots
dodge <- position_dodge(width = 0.6)
#Set species order for facet plotting
sp$Species=factor(sp$Species, levels=c('Zebra Finch', "House Sparrow", "E. Fence Lizard", "Brown Anole"))

fancy_scientific <- function(l) {
     # turn in to character string in scientific notation
     l <- format(l, scientific = TRUE)
     # quote the part before the exponent to keep all the digits
     l <- gsub("^(.*)e", "'\\1'e", l)
     # turn the 'e+' into plotmath format
     l <- gsub("e", "%*%10^", l)
     # return this as an expression
     parse(text=l)
}
```

#Facet plots of qPCR data
```{r}

overall.SQ=summarySE(sp, measurevar = "SQ", groupvars = c("Species","Individual", "Target", "Age"))

p1=ggplot(sp, aes(x=Age, y=SQ, fill= Target)) + 
    ylab("Relative Copy Number") +
  xlab("Life Stage") +
  ggtitle("IGF Relative Liver Expression") +
  geom_violin(trim=F, position= dodge, scale="width") +     
    scale_fill_manual(values = c("orange2", "skyblue4")) +
  geom_boxplot(width=0.15, position= dodge, outlier.shape = NA, color="black") +
geom_point(data = overall.SQ, size =2, shape = 19, color="black", position=position_dodge(width=0.6)) +
geom_point(position=position_jitterdodge(jitter.width = 0.05, dodge.width = 0.6), size=1, alpha=0.5, aes(group= Target), color="white") + 
     scale_y_continuous(labels=fancy_scientific) +
  theme_bw() +
  facet_wrap(~Species, scales="free", nrow=4) +
  theme(strip.text = element_text(size=14, face="bold"),
        axis.text.x = element_text(size=12)) 
p1

ggsave(p1, file="qPCR_facet.nc.png", width=6, height=9, dpi=600)

```

##anoles minus adults for inset
```{r}
BA=subset(sp, Species == "Brown Anole")
devan=subset(BA, Age != 'Adult')
devan.SQ=summarySE(devan, measurevar = "SQ", groupvars = c("Individual", "Target", "Age"))

p2=ggplot(devan, aes(x=Age, y=SQ, fill= Target)) + 
    ylab("Relative Copy Number") +
  xlab("Life Stage") +
  ggtitle("IGF Relative Liver Expression") +
  geom_violin(trim=F, position= dodge, scale="width") +     
    scale_fill_manual(values = c("orange2", "skyblue4")) +
  geom_boxplot(width=0.15, position= dodge, outlier.shape = NA, color="black") +
geom_point(data = devan.SQ, size =2, shape = 19, color="black", position=position_dodge(width=0.6)) +
geom_point(position=position_jitterdodge(jitter.width = 0.05, dodge.width = 0.6), size=1, alpha=0.5, aes(group= Target), color="white") + 
     scale_y_continuous(labels=fancy_scientific) +
  theme(axis.text.x = element_text(size=12))


p2

ggsave(p2, file="BA_dev_inset.nc.png", width=9, height=7, dpi=600)


```

#note: Run these again with SQ values
##Statistical Analysis
#ZebraFinch Statistics
```{r}
#Zebra Finch Statistics
sp=read.csv("Species_Combined_edited_2.csv")

zf=subset(sp, Species == "Zebra Finch")

zf.emb=subset(zf, Age=="Embryonic")
zf.juv=subset(zf, Age=="Juvenile")
zf.ad=subset(zf, Age=="Adult")

summary(lme(SQ~Target, random=~1|Content, data=zf.emb))
summary(lme(SQ~Target, random=~1|Content, data=zf.juv)) 
summary(lme(SQ~Target*Sex, random=~1|Content, data=zf.ad)) 

zf.ad.m=subset(zf.ad, Sex =="Male")
zf.ad.f=subset(zf.ad, Sex=="Female")
summary(lme(SQ~Target, random=~1|Content, data=zf.ad.m)) 
summary(lme(SQ~Target, random=~1|Content, data=zf.ad.f)) 

```

#House Sparrow Statistics
```{r}
#House Sparrow Statistics
hs=subset(sp, Species == "House Sparrow")

hs.juv=subset(hs, Age=="Juvenile")
hs.ad=subset(hs, Age=="Adult")

summary(lme(SQ~Target, random=~1|Content, data=hs.juv))
summary(lme(SQ~Target, random=~1|Content, data=hs.ad)) 

```

#Sceloperus Statistics
```{r}
#Scelop Statistics
scelop=subset(sp, Species == "E. Fence Lizard")

sc.juv=subset(scelop, Age=="Juvenile")
sc.ad=subset(scelop, Age=="Adult")

summary(lme(SQ~Target, random=~1|Content, data=sc.juv))
summary(lme(SQ~Target, random=~1|Content, data=sc.ad)) 

```
#BA Statistics
```{r}
#Anole Statistics
anole=subset(sp, Species == "Brown Anole")

a.emb=subset(anole, Age=="Embryonic")
a.juv=subset(anole, Age=="Juvenile")
a.ad=subset(anole, Age=="Adult")

summary(lme(SQ~Target, random=~1|Content, data=a.emb)) 
summary(lme(SQ~Target, random=~1|Content, data=a.juv)) 
summary(lme(SQ~Target*Sex, random=~1|Content, data=a.ad))

a.ad.m=subset(a.ad, Sex =="Male")
a.ad.f=subset(a.ad, Sex=="Female")
summary(lme(SQ~Target, random=~1|Content, data=a.ad.m)) 
summary(lme(SQ~Target, random=~1|Content, data=a.ad.f)) 

```
#Mouse Statistics
```{r}
#Mouse Statistics
mouse=subset(sp, Species == "House Mouse")

m.emb=subset(mouse, Age=="Embryonic")
m.juv=subset(mouse, Age=="Juvenile")
m.ad=subset(mouse, Age=="Adult")

summary(lme(SQ~Target, random=~1|Content, data=m.emb))
summary(lme(SQ~Target, random=~1|Content, data=m.juv)) 
summary(lme(SQ~Target*Sex, random=~1|Content, data=m.ad))

```

#Mouse Subset Data Plots by Strain (qPCR Data)
```{r}
mouse$Age_sub = factor(mouse$Age_sub, levels=c('Embryonic.Lab','Adult.Lab','Juvenlie.Wild', 'Adult.Wild'))

install.packages("ggforce")
install.packages("ggalt")
library(ggforce)
library(ggalt)
library(dplyr)

mean.SQ=summarySE(mouse, measurevar = "SQ", groupvars = c("Individual", "Target", "Age_sub"))


p5=ggplot(mouse, aes(x=Age_sub, y=SQ, fill= Target)) + 
    ylab("Relative Copy Number") +
  xlab("Life Stage") +
  ggtitle("IGF Relative Liver Expression") +
  geom_violin(trim=F, position= dodge, scale="width") +     
    scale_fill_manual(values = c("orange2", "skyblue4")) +
  geom_boxplot(width=0.15, position= dodge, outlier.shape = NA, color="black") +
    geom_point(data = mean.SQ, size =2, shape = 19, color="black", position=position_dodge(width=0.6)) +
geom_point(position=position_jitterdodge(jitter.width = 0.05, dodge.width = 0.6), size=1, alpha=0.5, aes(group= Target), color="white") + 
    theme_bw() +
  theme(axis.text.x = element_text(size=12)) 

p5

ggsave(p5, file="Mouse_Strain.nc.png", width=9, height=7, dpi=600)

```

#Mouse Subset Data Plots by Strain (RNAseq)
```{r}

#aggregate data to get average IGF1 for each species (combining all individuals of the same species)

IGF1.m=summarySE(data = data2, measurevar="IGF1.plot", groupvars = c("Strain", "Organism", "Category", "Subcategory", "Phylogenetic_Cat"), na.rm = FALSE, conf.interval = 0.95, .drop = TRUE)
IGF2.m=summarySE(data = data2, measurevar="IGF2.plot", groupvars = c("Strain", "Organism", "Category", "Subcategory", "Phylogenetic_Cat"), na.rm = FALSE, conf.interval = 0.95, .drop = TRUE)

dat=merge(x = IGF1.m, y = IGF2.m, by = "Strain", all.x=TRUE)

dat=data.frame(dat$Strain, dat$Organism.x ,dat$IGF1.plot, dat$IGF2.plot)
dat=dat[complete.cases(dat), ]

names(dat)[names(dat) == "dat.Strain"] <- "Strain"
names(dat)[names(dat) == "dat.IGF1.plot"] <- "IGF1"
names(dat)[names(dat) == "dat.IGF2.plot"] <- "IGF2"


dat3 <- dat %>% gather(key = "Gene", value="value", -c(1,2)) 

p6=ggplot(dat3, aes(fill=Gene, y=value, x=Strain)) +
  geom_bar(position="stack", stat="identity") +  scale_fill_manual(values=c("orange2", "skyblue4")) +
coord_flip() +
  scale_x_discrete(lim=rev) +
  theme(axis.title.y = element_blank(),
       legend.position = "top")

ggsave(p6, file="Mouse_Strain_RNAseq.nc.png", width=5, height=7, dpi=600)

```

#Graph of adult data by sex when available 
```{r}
sex=subset(sp, Age == "Adult"  & Species != "House Mouse" & Species != "E. Fence Lizard" & Species != "House Sparrow")
sex.SQ=summarySE(sex, measurevar = "SQ", groupvars = c("Individual","Species", "Sex", "Target"))


p7=ggplot(sex, aes(x=Sex, y=SQ, fill= Target)) + 
    ylab("Relative Copy Number") +
  xlab("Sex") +
  ggtitle("IGF Relative Liver Expression") +
  geom_violin(trim=F, position= dodge, scale="width") +     
    scale_fill_manual(values = c("orange2", "skyblue4")) +
  geom_boxplot(width=0.15, position= dodge, outlier.shape = NA, color="black") +
    geom_point(data = sex.SQ, size =2, shape = 19, color="black", position=position_dodge(width=0.6)) +
  geom_point(position=position_jitterdodge(jitter.width = 0.05, dodge.width = 0.6), size=1, alpha=0.5, aes(group= Target), color="white") + 
  theme_bw() +
  theme(axis.text.x = element_text(size=12)) +
  facet_wrap(~Species, scales="free") +
    theme(strip.text = element_text(size=14, face="bold")) 
  

p7

ggsave(p7, file="Sex_interactions.nc.png", width=9, height=5, dpi=600)

```

#Test for Correlations in IGF1 and IGF2 within species
```{r}
library(ggalt)
library(ggforce)

corr$Age = factor(corr$Age, levels=c('Embryonic','Juvenile', 'Adult'))


corr=read.csv("correlational_data.csv")
zf.cor=subset(corr, Species == "Zebra Finch")
ba.cor=subset(corr, Species == "Brown Anole")
hs.cor=subset(corr, Species == "House Sparrow")
sc.cor=subset(corr, Species == "E. Fence Lizard")

cor.test(zf.cor$IGF1, zf.cor$IGF2)

ggplot(corr, aes(IGF1, IGF2)) + 
  geom_point() + 
  geom_smooth(method="lm", color="black")+  
  geom_mark_ellipse(expand = 0,aes(fill=Age)) + 
  facet_wrap(~Species, scales="free")


ggplot(zf.cor, aes(IGF1, IGF2)) + geom_point() + geom_smooth(method="lm", color="black")+  geom_mark_ellipse(expand = 0,aes(fill=Age))
ggplot(ba.cor, aes(IGF1, IGF2)) + geom_point() + geom_smooth(method="lm", color="black")+  geom_mark_ellipse(expand = 0,aes(fill=Age))
ggplot(hs.cor, aes(IGF1, IGF2)) + geom_point() + geom_smooth(method="lm", color="black")+  geom_mark_ellipse(expand = 0,aes(fill=Age))
ggplot(sc.cor, aes(IGF1, IGF2)) + geom_point() + geom_smooth(method="lm", color="black")+  geom_mark_ellipse(expand = 0,aes(fill=Age))

```
